Community-based marine reserves produce biological and economic benefits

### 2018-02-08 Frontiers_Manuscript/Villasenor-Derbez.Rmd
updated manuscript

---
title: "Community-based marine reserves produce biological and economic benefits"
rtitle: "Mexican marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Stuart Fulton$^{2}$, Jorge Torre$^{2}$"
institution: "
$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: ""
keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
# biblio: test.bib
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
# csl: FMS.csl
---

@@

La sobrepesca y prácticas pesqueras no sostenibles son unas de las mayores amenazas para la conservación de los ecosistemas marinos del mundo \citep{halpern_2008-dK,halpern_2017-Zi}. La implementación de reservas marinas (*i.e.* áreas donde la captura de una o más especies está prohibida) es una medida de manejo frecuentemente propuesta para recuperar stocks pesqueros e impulsar la productividad pesquera en aguas cercanas \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recientes trabajos han demostrado que también pueden mitigar y proveer amortiguamiento ante el cambio climático \citep{roberts_2017-J9}, variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la biomasa, riqueza y densidades de organismos dentro de sus fronteras \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.

@@

# Load indicator functions
source(file = "../Scripts/abundance.R")
source(file = "../Scripts/richness.R")
source(file = "../Scripts/shannon.R")

@@

```{r load_data}
fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)"))

@@

invert_complete <- invert %>%
  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
  summarize(Abundancia = sum(Abundancia)) %>% 
  ungroup() %>% 
  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
  arrange()
```

@@

@@

F_H <- plot_model(model = list(H_IN, H_ME, H_PH),
                  various = T,
                  model_names = c("IN", "ME", "PH"))
# Invert plots
I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                  various = T,
                  model_names = c("IN", "ME", "PH"))

@@

>>>>

---
title: "Community-based marine reserves produce biological and economic benefits"
rtitle: "Mexican marine reserves"
author: "Juan Carlos Villaseñor-Derbez$^{1,*}$, Eréndira Aceves-Bueno$^{1,*}$, Stuart Fulton$^{2}$"
institution: "
$^{1}$Bren School of Environmental Science and Management, University of California, Santa Barbara, Santa Barbara, CA, USA\\newline
$^{2}$Comunidad y Biodiversidad A.C.,  Guaymas, Mexico
"
corrauth: "Juan Carlos Villaseñor-Derbez, Bren Hall, University of California, Santa Barbara, Santa Barbara, CA, 93106"
email: "jvillasenor@bren.ucsb.edu"
abstract: ""
keywords: "Marine Reserves, Marine Conservation, Small Scale Fisheries, Citizen Science, Mexico."
output:
  bookdown::pdf_document2:
    keep_tex: yes
    template: template.tex
---

@@

Overfishing and unsustainable fishing practices threaten marine ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}. The implementation of marine reserves (*i.e.* areas where all fishing activities are off--limits; MRs) is a common management approach used to manage the spatial distribution of fishing effort and recover stocks \citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recent research poses that marine reserves can also help buffer climate change \citep{roberts_2017-J9}, provide a refuge to environmental variability \citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm} and, in general, increase biomass, richness and densities of organisms within the protected regions \citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}. However, the extent to which MRs yield benefits to fisheries remains relatively unexplored \citep{krueck_2017-J1} and the focus is often only in the social and economic dimensions \citep{klein_2008,charles_2008}, rarely providing a hollistic evaluation of the social-ecological system \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. Here, we combine causal inference techniques \citep{depalma_2018} and the social-ecological systems framework \citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive evaluation of four community-based marine reserves in three coastal communities in Mexico.

@@

# Load indicator functions
source(file = "../Scripts/abundance.R")
source(file = "../Scripts/richness.R")
source(file = "../Scripts/shannon.R")
source(file = "../Scripts/biomass.R")
source(file = "../Scripts/trophic.R")

@@

```{r load_data}
fish <- read.csv("../Data/fish_data.csv", strip.white = T, stringsAsFactors = F) %>% 
  filter(Comunidad %in% c("Isla Natividad", "Maria Elena", "Punta Herrero")) %>% 
  filter(!is.na(RC),
         !RC %in% c("Punta Prieta - La Guanera", "El Gallinero - El Ganillero (Control)", "El Faro - El Faro (Control)", "Anegado de Chal - Anegado de Chal (Control)")) %>% 
  mutate(GeneroEspecie = case_when(
    GeneroEspecie == "Cephalopholis cruentatus" ~ "Cephalopholis cruentata",
    GeneroEspecie == "Haemulon macrostomun" ~ "Haemulon macrostomum",
    GeneroEspecie == "Xcochin spp" ~ "Balistes capriscus",
    GeneroEspecie == "Pterois volitans/miles" ~ "Pterois volitans",
    GeneroEspecie == "Mycteroperca spp" ~ "Mycteroperca",
    GeneroEspecie == "Chromis spp" ~ "Chromis",
    GeneroEspecie == "Damisela spp" ~ "Stegastes",
    GeneroEspecie == "Haemulon spp" ~ "Haemulon",
    GeneroEspecie == "Lancero spp" ~ "Acanthurus",
    GeneroEspecie == "Kyphosus sectatrix/incisor" ~ "Kyphosus",
    GeneroEspecie == "Sebastes spp" ~ "Sebastes",
    GeneroEspecie == "Loro spp" ~ "Scarus",
    GeneroEspecie == "Lutjanus spp" ~ "Lutjanus",
    GeneroEspecie == "Mariposa spp" ~ "Chaetodon",
    GeneroEspecie == "Pluma spp" ~ "Calamus pennatula",
    GeneroEspecie == "Sparisoma aurolineatum" ~ "Sparisoma aurofrenatum",
    GeneroEspecie == "Vieja spp" ~ "Bodianus rufus",
    # GeneroEspecie == "" ~ "",
    TRUE ~ GeneroEspecie
  ))

@@

data("species_bio")

spp <- group_by(fish_complete, GeneroEspecie) %>% 
  count() %>% 
  left_join(species_bio, by = "GeneroEspecie") %>% 
  filter(is.na(a) | is.na(b))

FB_update <- rfishbase::length_weight(c("Kyphosus incisor", "Kyphosus sectatrix", "Pterois volitans", "Scarus guacamaia", "Chaetodon capistratus", "Chaetodon capistratus", "Chaetodon striatus", "Calamus pennatula"), fields = c("a", "b")) %>% 
  select(GeneroEspecie = sciname, a, b) %>% 
  mutate(GeneroEspecie = case_when(
    GeneroEspecie %in% c("Kyphosus sectatrix", "Kyphosus incisor") ~ "Kyphosus",
    GeneroEspecie %in% c("Chaetodon capistratus", "Chaetodon striatus") ~ "Chaetodon",
    TRUE ~ GeneroEspecie
  )) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T)

GenusLevelBio <- species_bio %>% 
  filter(Genus %in% c("Mycteroperca", "Chromis", "Stegastes", "Haemulon", "Acanthurus", "Sebastes", "Lutjanus", "Scarus")) %>%
  select(GeneroEspecie = Genus, a, b) %>% 
  group_by(GeneroEspecie) %>% 
  summarize_all(mean, na.rm = T)

species_bio %<>%
  select(GeneroEspecie, a, b) %>% 
  rbind(FB_update) %>% 
  rbind(GenusLevelBio)
```

```{r complete_inverts}
invert_complete <- invert %>%
  select(Ano, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
  group_by(Ano, Sitio, Zona, Transecto, GeneroEspecie) %>% 
  summarize(Abundancia = sum(Abundancia)) %>% 
  ungroup() %>% 
  complete(Ano, Transecto, nesting(Sitio, Zona, GeneroEspecie), fill = list(Abundancia = 0L)) %>% 
  left_join(fish_covars, by = c("Ano", "Sitio", "Zona", "Transecto")) %>% 
  select(Comunidad, Ano, RC, Sitio, Zona, Transecto, GeneroEspecie, Abundancia) %>% 
  arrange()
```

@@

## Biomass
B_IN <- fish %>% 
  biomass(com = "Isla Natividad") %>% 
  model(sfe = F)

B_ME <- fish%>% 
  biomass(com = "Maria Elena") %>% 
  model(sfe = T)

B_PH <- fish %>% 
  biomass(com = "Punta Herrero") %>% 
  model(sfe = F)

## Trophic level
# T_IN <- fish %>% 
#   trophic("Isla Natividad") %>% 
#   model(sfe = F, spfe = F)
# 
# T_ME <- fish %>% 
#   trophic("Maria Elena") %>% 
#   model(spfe = F)
# 
# T_PH <- fish %>% 
#   trophic("Punta Herrero") %>% 
#   model(sfe = F, spfe = F)

@@

F_H <- plot_model(model = list(H_IN, H_ME, H_PH),
                  various = T,
                  model_names = c("IN", "ME", "PH"))

# Invert plots
I_N <- plot_model(model = list(Ni_IN, Ni_ME, Ni_PH),
                  various = T,
                  model_names = c("IN", "ME", "PH"))

@@

```{r, fig.width = 4, fig.height = 3}
F_B <- plot_model(model = list(B_IN, B_ME, B_PH),
                 various = T,
                 model_names = c("IS", "ME", "PH"))

F_B
```

### 2018-02-08 Frontiers_Manuscript/Villasenor-Derbez.tex
updated manuscript

\documentclass{frontiersSCNS}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}

@@

\usepackage[utf8]{inputenc}

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Stuart Fulton\(^{2}\), Jorge
Torre\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

La sobrepesca y prácticas pesqueras no sostenibles son unas de las
mayores amenazas para la conservación de los ecosistemas marinos del
mundo \citep{halpern_2008-dK,halpern_2017-Zi}. La implementación de
reservas marinas (\emph{i.e.} áreas donde la captura de una o más
especies está prohibida) es una medida de manejo frecuentemente
propuesta para recuperar stocks pesqueros e impulsar la productividad
pesquera en aguas cercanas
\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recientes
trabajos han demostrado que también pueden mitigar y proveer
amortiguamiento ante el cambio climático \citep{roberts_2017-J9},
variabilidad ambiental \citep{micheli_2012-EU}, resolver problemas de
pesca incidental \citep{hastings_2017-sm} y, en general, incrementar la
biomasa, riqueza y densidades de organismos dentro de sus fronteras
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}.

@@

>>>>

\documentclass{frontiersSCNS}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage[onehalfspacing]{setspace}
\usepackage{float}

@@

\usepackage[utf8]{inputenc}
\floatplacement{figure}{H}

\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Villaseñor-Derbez {et~al.}}
\def\Authors{Juan Carlos Villaseñor-Derbez\(^{1,*}\), Eréndira
Aceves-Bueno\(^{1,*}\), Stuart Fulton\(^{2}\)}
% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
\def\Address{\(^{1}\)Bren School of Environmental Science and Management, University
of California, Santa Barbara, Santa Barbara, CA,
USA\newline \(^{2}\)Comunidad y Biodiversidad A.C., Guaymas, Mexico}
% The Corresponding Author should be marked with an asterisk
% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
\def\corrAuthor{Juan Carlos Villaseñor-Derbez, Bren Hall, University of California,
Santa Barbara, Santa Barbara, CA, 93106}

@@

Overfishing and unsustainable fishing practices threaten marine
ecosystems around the world \citep{halpern_2008-dK,halpern_2017-Zi}. The
implementation of marine reserves (\emph{i.e.} areas where all fishing
activities are off--limits; MRs) is a common management approach used to
manage the spatial distribution of fishing effort and recover stocks
\citep{afflerbach_2014-HP,krueck_2017-J1,sala_2017-69}. Recent research
poses that marine reserves can also help buffer climate change
\citep{roberts_2017-J9}, provide a refuge to environmental variability
\citep{micheli_2012-EU}, help reduce by-catch \citep{hastings_2017-sm}
and, in general, increase biomass, richness and densities of organisms
within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}. However, the
extent to which MRs yield benefits to fisheries remains relatively
unexplored \citep{krueck_2017-J1} and the focus is often only in the
social and economic dimensions \citep{klein_2008,charles_2008}, rarely
providing a hollistic evaluation of the social-ecological system
\citep{halpern_2013,lpezangarita_2014,mascia_2017-m_}. Here, we combine
causal inference techniques \citep{depalma_2018} and the
social-ecological systems framework
\citep{ostrom_2009-hg,mascia_2017-m_} to provide a comprehensive
evaluation of four community-based marine reserves in three coastal
communities in Mexico.

@@

\includegraphics{Villasenor-Derbez_files/figure-latex/unnamed-chunk-4-1.pdf}
